---
title: "Degradation Functions in DA & NE"
date: "2020"
output: html_document
---

## Experimental Branch; create 'mesolimbic'

Using this rnotebook to try ideas for the ADHD ABM;object oriented abm using reference classess (http://adv-r.had.co.nz/R5.html).

```{r}

# Dopamine agent; makes DA: 

DA <- setRefClass("dopamine",
                  #Fields indicate attributes this reference class will have
                  fields = list(Tyrosine_hydroxylase = "numeric",
                                DOPA_decarboxylase = "numeric"),
                  
                  #methods indicate what functions the reference class can perform
                  methods = list(
                    
                    #f(x) 
                    Biosynthesize = function(tyrosine){
                      L_DOPA <<- tyrosine*Tyrosine_hydroxylase # how do we know how tyrosine to start with? 
                      Dopamine <<- L_DOPA*DOPA_decarboxylase
                      return(list(L_DOPA,
                                  Dopamine))
                      }
                    )
                  )
```

```{r}

# Norepinephrine agent; makes NE 

NE <- setRefClass("norepinephrine",
                  #Fields indicate attributes this reference class will have
                  fields = list(Tyrosine_hydroxylase = "numeric",
                                DOPA_decarboxylase = "numeric",
                                DOPA_Bhydroxylase = "numeric"),
                  
                  #methods indicate what functions the reference class can perform
                  methods = list(
                    
                    #f(x) 
                    Biosynthesize = function(tyrosine){
                      L_DOPA <<- tyrosine*Tyrosine_hydroxylase # how do we know how tyrosine to start with? 
                      Dopamine <<- L_DOPA*DOPA_decarboxylase
                      Norepinephrine <<- Dopamine*DOPA_Bhydroxylase
                      return(list(L_DOPA,
                                  Norepinephrine))
                      }
                    )
                  )
```


```{r}

# Norepinephrine agent; makes DA: 1) digests MPH tablet w/ aid of CESA1A & 2) transfering availed compounds to the bloodstream

ML <- setRefClass("mesolimbic",
                  #Fields indicate attributes this reerence class will have
                  fields = list(COMT = "numeric", # gets set by polymorph_activity
                                #polymorph_activity = "numeric",
                                MAO = "numeric",
                                ADH = "numeric",
                                Dopamine = "numeric"), #1/4 of upstream step
                  
                  #methods indicate what functions the reference class can perform
                  methods = list(
                    
                    #f(x) 
                    Degrade = function(Dopamine){ #include 1//4 splt again
                      
                      MT3 <<- Dopamine*COMT
                      MHPA <<- MT3*MAO
                      HVA <<- MHPA*ADH
                      
                      DHPA <<- Dopamine*MAO
                      DOPAC <<- MT3*ADH
                      HVA <<- MHPA*COMT
                      
                      return(list(COMT,
                                  MAO,
                                  ADH,
                                  MT3,
                                  MHPA,
                                  HVA,
                                  DHPA,
                                  DOPAC))
                      }
                    )
                  )
```

```{r}


#liver starts off empty with no naturally occuring D_MPH or L_MPH, and some specified level of CES1A1 ??
ml_0 <- ML$new(COMT = 0.8,
               MAO = 0.6,
               ADH = 0.5,
               Dopamine = 250)

#provide gut with a tablet to 'digest'; tablet D_MPH and L_MPH need to be specified 
ml_0$Degrade(100) 

```

```{r}
for (i in 1:100){
  degraded_product <- ml_0$Degrade(100)[[6]]
  print(degraded_product)
} 

```

```{r}
plot(degraded_product)

```



```{r}


#liver starts off empty with no naturally occuring D_MPH or L_MPH, and some specified level of CES1A1 ??
da_0 <- DA$new(Tyrosine_hydroxylase=0.8,
               DOPA_decarboxylase = 0.6)
#provide gut with a tablet to 'digest'; tablet D_MPH and L_MPH need to be specified 
da_0$Biosynthesize(100) 

```

```{r}
library(tidyverse)



enzyme_df <- read_csv("Copy of CES1_ADHDME_kids_LIVER_score.csv")

#Standardize CES1A1 PRS
enzyme_df$PRS_1.0 <- (enzyme_df$PRS_1.0 - min(enzyme_df$PRS_1.0)) / (max(enzyme_df$PRS_1.0) - min(enzyme_df$PRS_1.0))
enzyme_df$ID <- as.numeric(str_sub(enzyme_df$ID1, start = -4))

local_enzyme <- enzyme_df%>% filter(ID == 003) ## SAMPLE ID PROVIDED TO AGENT 
ind_Tyrosine_hydroxylase <- local_enzyme$PRS_1.0
ind_DOPA_decaroxylas <- local_enzyme$PRS_1.0

#liver starts off empty with no naturally occuring D_MPH or L_MPH, and some specified level of CES1A1 ??
da_1 <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
               DOPA_decarboxylase = ind_DOPA_decaroxylas)
#provide gut with a tablet to 'digest'; tablet D_MPH and L_MPH need to be specified 
da_1$Biosynthesize(100) 
```


```{r}
# Dopa neuron (subtantia niagra to neostratum)
da_sn_ns <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                   DOPA_decarboxylase = ind_DOPA_decaroxylas)

# Dopa neuron (vta to nucleus acumbens)
da_vta_na <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                    DOPA_decarboxylase = ind_DOPA_decaroxylas)

# Dopa neuron (vta to pfc); needs adapting to include DA - D2 & DA - DA (***)
da_vta_pfc <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                    DOPA_decarboxylase = ind_DOPA_decaroxylas)

da_vta_pfc_d1 <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                        DOPA_decarboxylase = ind_DOPA_decaroxylas)

da_vta_pfc_d2 <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                        DOPA_decarboxylase = ind_DOPA_decaroxylas)

da_vta_pfc_d3 <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                        DOPA_decarboxylase = ind_DOPA_decaroxylas)

da_vta_pfc_d4 <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                        DOPA_decarboxylase = ind_DOPA_decaroxylas)

# Dopa neuron (hypothalamus)
da_hypothalamus <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                    DOPA_decarboxylase = ind_DOPA_decaroxylas)

da_hypothalamus_d1 <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                        DOPA_decarboxylase = ind_DOPA_decaroxylas)

da_hypothalamus_d2 <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                        DOPA_decarboxylase = ind_DOPA_decaroxylas)

da_hypothalamus_d3 <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                        DOPA_decarboxylase = ind_DOPA_decaroxylas)

da_hypothalamus_d4 <- DA$new(Tyrosine_hydroxylase = ind_Tyrosine_hydroxylase,
                        DOPA_decarboxylase = ind_DOPA_decaroxylas)
```

Example of for loop for calling any one of the above: 
```{r}
for (i in 1:100){
  L_DOPA <- da_vta_na$Biosynthesize(100)[[1]]
  Dopamine <- da_vta_na$Biosynthesize(100)[[2]]
} 

```

For each of the above: 
1) we can call it x times (or based on another variable) to show relative activation (we may have said 0.25 for each?). This would show how brain connections are networks that we explictly represent in our agent based model; 
2) Input for biosynthesis would need to based on outputs of upstream steps (right now we have as.numeric, will need attention later / now?)
3) Do we need baseline activation (as is the case irl)
4) Degradation -- do we need a new global function for this? Or do we want degradation to be based on upstream biosynthesis?