---
title: "R Notebook"
output: html_notebook_none
---

## More explicit control structure

Adding an explicit network structure to the agents, and a initialization phase (parameterize for each individual)

## 1. Initialization phase: 

1) import individual-level parameters from genetic database. These parameters will be used when creating the various agents that make up the ABM.


import individual parameters for the ABM
```{r}

#which individual should we run the simulation for
ID = 5

#bring in the data and extract individual parameters... simple for now:
df.ces1a1 <- read.csv("data/Data file_July172020.csv")

#nromalize the values until we can figure out why some are negative...
df.ces1a1$norm_ces1a1 <- (df.ces1a1$CES1A1 - min(df.ces1a1$CES1A1)) / (max(df.ces1a1$CES1A1) - min(df.ces1a1$CES1A1))

#set the individuals parameters
ind_CES1A1=df.ces1a1$norm_ces1a1[ID]


#input pill parameters
D_MPH = 1000
L_MPH = 1000
```


## 2. Define the various agents

Define all agents following a network format. All agents have a step function that i) uses information from all incomming edges, ii) updates the agent, and iii) outputs all information/properties to outgoing edges. (might have to seperate these functions if we want to run this all synchronously...)
```{r}

#example of a node/agent
agent_gut <- setRefClass("gut",
                         
                         #what state variables does this agent have
                         fields = list(tablet_D_MPH = "numeric",
                                       tablet_L_MPH = "numeric",
                                       in_edges = "list",
                                       out_edges = "list"),
                         
                         #what functions does this agent have
                         methods = list(
                           
                           step = function(net){
                            
                             #edges to modify dataframe
                             df.modify <- list()
                             
                             #1. update agent properties
                             #nothing (no in edges)
                             
                             #2. step agent
                             conv <- digest_tablet()
                             
                             #3. pass along info/properties
                             for(e in out_edges){
                               
                               df.modify[[1]] <- list(from=e[2],to=e[3],type="D_MPH",weight=conv$D_conv)
                               df.modify[[2]] <- list(from=e[2],to=e[3],type="L_MPH",weight=conv$L_conv)
                               
                               #E(net,P=e[2:3])$D_MPH <- conv$D_conv
                               #E(net,P=e[2:3])$L_MPH <- conv$L_conv
                               
                               #print(paste0("gut passes to blood ",E(net,P=e[2:3])$L_MPH," L_MPH"))
                             }
                             
                             #return the modified network
                             return(df.modify)
                             
                           },
                           
                           #Function to digest tablet
                           digest_tablet = function(){
                             
                             #convert from tablet to gut
                             D_conv <- tablet_D_MPH*0.10 #this needs to be fixed
                             L_conv <- tablet_L_MPH*0.10 #this needs to be fixed
                             tablet_D_MPH <<- tablet_D_MPH - D_conv
                             tablet_L_MPH <<- tablet_L_MPH - L_conv
                             
                             return(list(D_conv=D_conv,L_conv=L_conv))
                           }, 
                             
                             add_in_edge = function(edge){
                               in_edges <<- append(in_edges,edge)
                             },
                             
                             add_out_edge = function(edge){
                               out_edges <<- append(out_edges,edge)
                             } #might have to place functions to remove edges if connections get dynamic...
                           
                         )
)

agent_blood <- setRefClass("blood",
                           
                           #what state variables does this agent have
                           fields = list(D_MPH = "numeric",
                                         L_MPH = "numeric",
                                         in_edges = "list",
                                         out_edges = "list"),
                           
                           #what functions does this agent have
                           methods = list(
                             
                             step = function(net){
                               
                               #edges to modify dataframe
                               df.modify <- list()
                               
                               #1. update agent properties
                               for(e in in_edges){
                                 
                                 #gut incoming MPH
                                 if(e[1] == "gut"){
                                   
                                   D_MPH <<- D_MPH + E(net,P=e[2:3])$D_MPH
                                   L_MPH <<- L_MPH + E(net,P=e[2:3])$L_MPH
                                   
                                   #print(paste0("blood takes ", E(net,P=e[2:3])$L_MPH, " L_MPH from the gut, now with ",L_MPH," L_MPH total"))
                                 }
                                 
                                 #liver incomming CES1A1 enzyme
                                 if(e[1] == "liver"){
                                   
                                   Amt_D_removed <- D_MPH*E(net,P=e[2:3])$CES1A1 
                                   Amt_L_removed <- L_MPH*E(net,P=e[2:3])$CES1A1
                                   D_MPH <<- D_MPH - Amt_D_removed
                                   L_MPH <<- L_MPH - Amt_L_removed
                                   
                                   #print(paste0("blood takes",E(net,P=e[2:3])$CES1A1," CES1A1 from the liver and removes ",Amt_L_removed," L_MPH, leaving ", L_MPH, " in the blood") )
                                   
                                   
                                 }
                                 
                               }
                               
                               #2. Step the agent
                               #nothing
                               
                               #3. pass info/properties to neighbouring agents
                               for(e in out_edges){
                                 
                                 df.modify[[1]] <- list(from=e[2],to=e[3],type="D_MPH",weight=D_MPH)
                                 df.modify[[2]] <- list(from=e[2],to=e[3],type="L_MPH",weight=L_MPH)
                               
                                 #E(net,P=e[2:3])$D_MPH <- D_MPH
                                 #E(net,P=e[2:3])$L_MPH <- L_MPH
                                 
                                 #print(paste0("blood passes L_MPH to the brain: ",E(net,P=e[2:3])$L_MPH))
                               }
                               
                               
                             #return the modified network
                             return(df.modify)
                               
                               
                             }, 
                             
                             add_in_edge = function(edge){
                               in_edges <<- append(in_edges,edge)
                             },
                             
                             add_out_edge = function(edge){
                               out_edges <<- append(out_edges,edge)
                             } #might have to place functions to remove edges if connections get dynamic...
                             
                           )
)

#create a simple agent that takes two inputs, and has a function that returns nothing.
agent_liver <- setRefClass("liver",
                           
                           #what state variables does this agent have
                           fields = list(CES1A1 = "numeric",
                                         in_edges = "list",
                                         out_edges = "list"),
                           
                           #what functions does this agent have
                           methods = list(
                             
                             step = function(net){
                               
                               #edges to modify
                               df.modify <- list()
                               
                               #1. update agent properties
                               #nothing
                               
                               #2. Step the agent
                               #nothing
                               
                               #3. pass info/properties to neighbouring agents
                               for(e in out_edges){

                                 #E(net,P=e[2:3])$CES1A1 <- CES1A1
                                 df.modify[[1]]<-list(from=e[2],to=e[3],type="CES1A1",weight=CES1A1)
                                 
                                 
                                 #print(paste0("liver passes CES1A1 to the blood: ",E(net,P=e[2:3])$CES1A1))
                               }
                               
                               
                             #return the modified network
                             return(df.modify)
                               
                               
                             }, 
                             
                             add_in_edge = function(edge){
                               in_edges <<- append(in_edges,edge)
                             },
                             
                             add_out_edge = function(edge){
                               out_edges <<- append(out_edges,edge)
                             } #might have to place functions to remove edges if connections get dynamic...
                             
                           )
)


#create a simple agent that takes two inputs, and has a function that returns nothing.
agent_dopaminergic_neuron <- setRefClass("Dopa",
                           
                           #what state variables does this agent have
                           fields = list(D_MPH = "numeric",
                                         L_MPH = "numeric",
                                         DA = "numeric",
                                         ldopa = "numeric",
                                         tyrosine = "numeric",
                                         vesicles = "numeric",
                                         degrade_on_uptake = "numeric",
                                         in_edges = "list",
                                         out_edges = "list"),
                           
                           #what functions does this agent have
                           methods = list(
                             
                             step = function(net){
                               
                               #edges to modify
                               df.modify <- list()
                               
                               #1. update agent properties
                               for(e in in_edges){
                                 
                                 #blood incoming MPH
                                 if(e[1] == "blood"){
                                   
                                   D_MPH <<- E(net,P=e[2:3])$D_MPH
                                   L_MPH <<- E(net,P=e[2:3])$L_MPH
                                   
                                   #print(paste0("Dopa Neuron is exposed to ", E(net,P=e[2:3])$L_MPH, " L_MPH from the blood) )
                                 }
                                 
                                 #inputs from other dopa neurons??
                                 if(e[1] == "Dopa"){
                                   
                                   #TO DO
                                   DA <<- DA + E(net,P=e[2:3])$DA #or is it the vesicles ... what should these agents pass to each other
                                   #print(paste0("Dopa neuron recives DA from other neurons: ",E(net,P=e[2:3])$DA))
                                   
                                 }
                                 
                               }
                               
                               
                               #2. Step the agent
                               tyrosine_to_lDOPA()
                               ldopa_to_DA()
                               DA_to_vesicles()
                               reuptake()
                               
                               
                               
                               #3. pass info/properties to neighbouring agents
                               for(e in out_edges){

                                 df.modify[[1]]<-list(from=e[2],to=e[3],type="DA",weight=DA) #or is it the vesicles ... what should these agents pass to each other
                                 
                                 #print(paste0("Dopa neuron passes DA to other neurons: ",E(net,P=e[2:3])$DA))
                               }
                               
                               
                             #return the modified network
                             return(df.modify)
                               
                               
                             }, 
                             
                             tyrosine_to_lDOPA = function(){
                               
                               #TO DO
                               
                             }, 
                             
                             ldopa_to_DA = function(){
                               
                               #TO DO
                               
                             },
                             
                             DA_to_vesicles = function(){
                               
                               #TO DO
                               
                             },
                             
                             reuptake = function(){
                               
                               #TO DO
                               
                             },
                             
                             add_in_edge = function(edge){
                               in_edges <<- append(in_edges,edge)
                             },
                             
                             add_out_edge = function(edge){
                               out_edges <<- append(out_edges,edge)
                             } #might have to place functions to remove edges if connections get dynamic...
                             
                           )
)


```


## 2. Creating the agent based model (network)

Create the agents and link them together to create the ABM.

```{r}

library(igraph)

#create an empty network/graph. 
system_graph <- make_empty_graph(directed = T)
all_agent_list <- list()
agent_index <- 1

#####################################
#
# build gut to brain system
#
#####################################

###################
### Create the gut
###################

gut <- agent_gut$new(tablet_D_MPH = 200, tablet_L_MPH = 100, out_edges=list() )
all_agent_list <- append(all_agent_list,gut)

# add the gut to the network
system_graph <- add.vertices(system_graph, agent_index, attr=c(name="gut", index =  agent_index))

#shift the index
agent_index = agent_index + 1

#######################
### Create the blood 
#######################

blood <- agent_blood$new(D_MPH = 0, L_MPH = 0, in_edges=list(), out_edges=list() )
all_agent_list <- append(all_agent_list,blood)

# add the blood to the network
system_graph <- add.vertices(system_graph, 1, attr=c(name="blood", index =  agent_index))

#shift the index
agent_index = agent_index + 1

#add in_edge in the network
system_graph <- add.edges(system_graph, c("gut","blood"), attr =list("type"="gut","D_MPH"=0,"L_MPH"=0)  )

#Tell the gut to output to the blood
gut$add_out_edge(list(info=c("gut","gut","blood")))

#let the blood listen to the edge from the gut
blood$add_in_edge(list(info=c("gut","gut","blood")))

######################
### Create the liver 
######################

liver <- agent_liver$new(CES1A1=ind_CES1A1, in_edges=list(), out_edges=list() )
all_agent_list <- append(all_agent_list,liver)

# add the blood to the network
system_graph <- add.vertices(system_graph, 1, attr=c(name="liver", index =  agent_index))

#shift the index
agent_index = agent_index + 1

#add edge in the network
system_graph <- add.edges(system_graph, c("liver","blood"), attr =list("type"="liver","CES1A1"=ind_CES1A1)  )

#Tel the liver to output to the blood
liver$add_out_edge(list(info=c("liver","liver","blood")))

#let the blood listen to the edge from the liver
blood$add_in_edge(list(info=c("liver","liver","blood")))   #format:: type, from , to


##########################
###add first dopa agent
##########################

dopa1 <- agent_dopaminergic_neuron$new(L_MPH=0,D_MPH=0,DA=0, in_edges=list(), out_edges=list() )
all_agent_list <- append(all_agent_list,dopa1)

# add the dopa agent to the network
system_graph <- add.vertices(system_graph, 1, attr=c(name="dopa1", index =  agent_index))

#shift the index
agent_index = agent_index + 1

#add edge in the network
system_graph <- add.edges(system_graph, c("blood","dopa1"), attr =list("type"="blood","D_MPH"=0,"L_MPH"=0)  ) #again not sure what to put here

#Let dopa1 listen to edge from blood
dopa1$add_in_edge(list(info=c("blood","blood","dopa1")))   #format:: type, from , to

############################
###add second dopa agent
############################

dopa2 <- agent_dopaminergic_neuron$new(L_MPH=0,D_MPH=0,DA=0, in_edges=list(), out_edges=list() )
all_agent_list <- append(all_agent_list,dopa2)

# add the dopa agent to the network
system_graph <- add.vertices(system_graph, 1, attr=c(name="dopa2", index =  agent_index))

#shift the index
agent_index = agent_index + 1

#add edge in the network
system_graph <- add.edges(system_graph, c("dopa1","dopa2"), attr =list("type"="dopa","DA"=0)  ) #again not sure what to put here

#Tell dopa1 to output to dopa2
dopa1$add_out_edge(list(info=c("dopa","dopa1","dopa2")))   #format:: type, from , to

#Let dopa2 lisen to in-edge from dopa 1
dopa2$add_in_edge(list(info=c("dopa","dopa1","dopa2")))   #format:: type, from , to


####################################
#
# Take a look at the ABM pathways
#
####################################

#take a look a the ABM 
plot(system_graph, edge.width = 2, vertex.size = 35 )

```


## 4. Run the agent based model

Use the network framework to control and update the abm (synchronous or asynchonous...)
```{r}
#lets do random asynchronous to start

D_MPH_in_blood <- vector()
D_MPH_in_DA2 <- vector()

#run 100 steps
for(i in 1:100){
  
  #################
  # Run all nodes #
  #################
  
  #process each node/agent (synchronously... i.e., order doesn't matter...)
  df.all.edges.to.modify <- list()
  for(n in 1:vcount(system_graph) ){
    df.all.edges.to.modify <- append(df.all.edges.to.modify,all_agent_list[n][[1]]$step(system_graph))
  }
  
  ######################
  # Update the network #
  ######################
  
  #update the network edges (synchronously... i.e., order doesn't matter)
  for(e in 1:length(df.all.edges.to.modify)){
    
    #get the edge in question
    e.temp <- df.all.edges.to.modify[[e]]
    
    #update the edge weight
    type = e.temp$type
    system_graph<-set_edge_attr(system_graph, name=type, index = E(system_graph,P=c(e.temp$from,e.temp$to)), value = e.temp$weight)
  }
  
  ######################
  # Record outputs     #
  ######################
  
  #record D_MPH in the blood
  D_MPH_in_blood[length(D_MPH_in_blood)+1] <- all_agent_list[[2]]$D_MPH
  
  #record DA in the dopa2 agent
  D_MPH_in_blood[length(D_MPH_in_blood)+1] <- all_agent_list[[5]]$D_MPH
  
}

#take a look at D_MPH in the blood
plot(D_MPH_in_blood)
```



